Using \LaTeX\ to submit manuscripts to peer reviewed journals.

### 2014-03-10 docs/thesis/chap-cern_math_explorer.tex
wrote clusring results work on document
proofreading fixes

\section{Features Extraction}
Our main goal is to develop a system that compliments the already available search capabilities in Invenio and in particular the CDS instance. Our system was developed on top of the  the Apache Solr/Lucene framework. CDS already uses Solr as a search engine layer, so developing $5e^{x+y}$(CDS) using this technology supposed a reasonable choice. One of the main steps while building an IR tool, is understanding the data and what kind of features can be extracted from it: If the system is based on text, then tasks as stemming, handling of synonyms, misspelling and so on, such be taken into consideration; if the system being developed is based on images, then examples of features can be histograms of colors, geometrical features and so on. During the literature review, we discovered that most of the projects handled the notational aspects of the expressions, that is, the real naming of the terms in a given equation and each project developed its own set of heuristics to handle the structural component.

\subsection{Notational Features}
Notational features correspond to the elements that relate to the actual naming of the variables, constants and numeric quantities found in an equation. Here we extract tokens that represent single leaf nodes (E.g. <mi>, <mn>, <mo> tags), and simple constructs like subscript and superscript elements, fractions, roots.
We also extract bi-grams and tri-grams as analogue to that is done for text indexing in current systems.

@@

\subsubsection{Unicode Normalizing} 
The first issue we discovered is that because of the big wide of different characters to represent the possible mathematical symbols, there is need for an automatic way to match different possible variations of a single character. An example of this situation is specification of an natural number, sometimes denoted by the letter N. In this case, at least three different common representations were identified: $ N = 1 $ (Using character with unicode code point 0x004e), $\mathcal{N} = 1$ (0xd835 0xdc41 or in latex expressed as \lstinline|\mathcal{N}|) and $\mathbb{N} = 1$ (0x2115 or in latex expressed as \lstinline|\mathbb{N}|). \\
  For addressing this situation, we rely on the Unicode equivalence framework. In the unicode specification, sometimes the same character have two or more different code-points because of backwards compatibility with other encoding standards or because the same character has different essential meanings. Another group of equal characters with different code-points are pre-combined characters such as the Latin letter \~{N} which has Unicode point 0x00D1 and also is the sequence of code-points  0x004E (Latin letter N) and 0x0303 (the combining character tilde). Taking this sample into account, Unicode defines the character composition and decomposition operations as replacing the decomposed representation with the pre-composed one and vice-versa. Finally, Unicode defines two types of equivalences between characters: Canonical and compatible. Canonical equivalent characters are assumed to have the same appearance and meaning. Compatible ones are allowed to have slightly different graphical representations but the same meaning in some contexts. One example of compatible equivalent characters is the Roman number 12 {\unicodefontⅫ} with code-point 0x216b and the sequence of characters XII with codepoints 0x0058 0x0049 0x0049. Taking into account the composition or decomposition of characters and the canonical or compatible equivalences, Unicode specifies 4 different forms of a given character:
  \begin{itemize}
  \item NFD: Normalization Form Canonical Decomposition
  \item NFC: Normalization Form Canonical Composition
  \item NFKD: Normalization Form Compatibility Decomposition
  \item NFKC: Normalization Form Compatibility Composition
  \end{itemize}

@@

Since our goal is to match as much as possible similar characters, our system employs the NFKD representation of a character. For a given token, we test whether if it is already equal to its NFKD. If not, for all the characters that are not combining nor control characters, we add the token into the same position as the original one.
For example, given the token <mi>\r{A}</mi>, this will lead to the sequence <mi>\r{A}</mi><mi>A</mi> and the token <mi>{\unicodefontⅫ}</mi> will produce <mi>{\unicodefontⅫ}</mi><mi>X</mi><mi>I</mi><mi>I</mi>.

\subsubsection{Operator grouping}
Even though the Unicode normalization step work for a big part of the cases presented in table \ref{unicode_characters}, there are still some groups of characters which would not be matched (Even partially), such as some types of integrals. For addressing these groups, a precomputed table was created for some general groups of operators which have some similar meaning. For each token that belongs to some of this predefined groups, a second token is created identifying the group.
For example the token <mo>{\unicodefont ∮}</mo> is expanded into <mo>{\unicodefont ∮}</mo><mo>INTEGRALS</mo>, similarly the token token <mo>{\unicodefont ∫}</mo> is expanded into <mo>{\unicodefont ∫}</mo><mo>INTEGRALS</mo> and then both will have a partial match in the INTEGRALS token. 
Unicode specification does not define a strict rule to group similar characters based on their semantics. For general operators, we used the grouping provided by Xah Lee in his website\cite{math_groups}.

@@

\subsubsection{Numeric Approximation} 
Some equations relate specific numeric quantities. While most of the previously explorer systems, handled <mn> tags by only representing them as a constant identifier, this approach is not suitable. A lot of articles in CDS relate to specific experiments under certain conditions, and require a high level of precision in the measurements. However, different experiments of the same phenomenon, can result in different measurements and we wanted to include this variability in our system. To handle this, when a token with a certain numeric quantity is discovered, we apply iterative rounding and index all the different approximations in the same position as the original element. For example the token <mn>2.0135</mn> will produce the following sequence of tokens: <mn>2.0135</mn><mn>2.014</mn><mn>2.01</mn><mn>2.0</mn>.

\subsection{Structural Features}
The structure of a mathematical expression compliments the information that an be obtained over generic substructures where the naming of the variables is irrelevant.

\subsubsection{Algebraic structures}
The first set of structures consist on purely algebraic relationships between the expressions.
Inspired by predefined integral tables for different kinds of expressions, we decided to extract from a given equation, whether it has the occurrence of these kind of patterns.
An example of this type of structures would be: $X\textunderscore*(X\textunderscore + A\textunderscore)$ where $X\textunderscore$ and $A\textunderscore$ can be any type of expression. Sample equations matching this pattern can be $C (C+1)$ or $\frac{(\sin{\theta}\cos{\theta})^2}{2a} (\frac{(\sin{\theta}\cos{\theta})^2}{2a} + \tan{\frac{\theta+\phi}{2}})$. This structural analysis of equations is more powerful than the unification of variables and constants proposed in previously discussed projects because these patterns, as shown in the second example, can match an arbitrary complex expression. The drawback of our solution is the need to manually select representative and useful patterns. This selection was done based on standard based tables of integrals similar to the ones that CASs employ to compute them.

\subsubsection{Subscript/Superscript variations}
This set of structures represent common operations that cat appear in a given expression, but its identification is based on notation factors. A simple example of this can be the matrix decomposition $A = VDV^{-1}$ where the actual name of the variables $A$ and $V$ is not meaningful but the real importance is the fact that a given variable appears alongside its inverse in a matrix multiplication form. We would like to assign a partial match to any other kind of this decomposition say $B = MRM^{-1}$. these types of structures live in-between the notational and structural categories but since we identify them using Mathematica, we encompass them as structural features. For this type of structures, we identify simple operations between an element and some possible variations of the same element with some subscript/superscript.

@@

Additionally, there are common patterns to specific domains which can improve the quality of the retrieved results. Some specific patterns we identified include: $originalParticle \rightarrow subparticle1, subparticle2, subparticle3$ (For representing particle decaying processes) $S = \int^{t2}_{t1} \mathcal{L} dt$ (For an action in terms of its Lagrangian) and $SU(n)$ (For describing special unitary groups). The main drawback of this approach is the need to manually detect, implement and maintain long lists of patterns and for this reason we did not tried to do an exhaustive list of them.

@@

Lucene framework defines some key concepts which ultimately are mapped into concrete classes. A {\codefont Document} is the minimum retrieval unit. For $5e^{x+y}$ we could either use a complete article as a document or a single equation. We choose the latter because it makes easier to focus on extracting features from a single equation. A document contains a set of {\codefont fields} which represent different types of information that it can store. Each field has different options such as whether it would be indexed and/or stored. A field can also indicate whether it will be a {\codefont Term Vector}. Our fields are as follow: 
\begin{itemize}
\item math\_notational\_field: This field stores the notational features as a Term Vector. 
\item math\_structural\_field: This field stores the structural features as a Term Vector. 
\item math\_normalized\_notational\_field: This field stores the notational features of the normalized expression as a Term Vector. 
\item filename: This is an auxiliary field which stores the CDS record / article name where the equation was found.
\item number\_occurrences: Another auxiliary field that is used to score final results.
\end{itemize}

Each field is associated to a particular {\codefont Analyzer}. An analyzer is a specific class that handles an element from a field and groups all the operations that are to be performed. For this purpose we wrote {\codefont SolrNotationalAnalyzer}, {\codefont SolrStructuralAnalyzer} and {\codefont SolrNormalizedNotationalAnalyzer}. 
The Solr prefix is added to make clear the fact that the class extends a {\codefont SolrAnalyzer} which extends a normal lucene {\codefont Analyzer}.

@@

\begin{itemize}
\item {\codefont MultiplePatternTokenizer}: This class is in charge of applying a set of regular expressions over the given expression in order to extract the single elements that were presented. Lucene incorporates the class {\codefont PatternTokenizer} but it only accepts one regular expression.
\item {\codefont UnicodeNormalizingFilter}: This class receives single tokens and verifies if the token is in Unicode NFKD. In negative case, it emits an additional token in the normalized form.
\item {\codefont SynonymExpandingFilter}: This class verifies if the received token belongs to a predefined category of elements and emits an additional token with it.
\item {\codefont NumericalRoundingFilter}: This class process {\codefont <mn>} tokens. When a number contains a decimal part, it applies the standard rounding mode iteratively and for each one, it emits a token for each step. 
\end{itemize}

@@

\subsubsection{Error handling}
Sometimes, after certain time, the connection between {\codefont KernelLink} and {\codefont Mathematica} will be lost launch an exception and following invocations would fails as well.
Also, re-instantiating a KernelLink will yield an Exception. For this scenario, killing the associated processes to Mathematica was necessary with {\codefont killall -s 9 Mathematica; killall -s 9 MathKernel
} or {\codefont taskkill /im Mathematica.exe; taskkill /im MathKernel.exe} depending on the OS the system is running on.

@@

Mathematica took around 10 minutes to interpret it.
In the ideal case we would be able to identify the malformed expressions and pass to Mathematica only well-formed equations to be processed but as this simple example shows, that is a very challenging task and may require more work.
As a remedial situation, we used a fixed timeout to stop the processing of a given expression.

@@

A demo interface over the {\codefont pu} branch was developed to showcase the system, however to have a complete integration with CDS, a more deep work is required. The demo setup employs the same dataset as used for experimentation.

\begin{figure}
\includegraphics[height=8 cm]{figures/input_interface.png}
\label{input_interface}
\caption{Web interface of $5e^{x+y}$: Query input screen}
\end{figure}

\begin{figure}
\includegraphics[height=8 cm]{figures/results_interface.png}
\label{results_interface}
\caption{Web interface of $5e^{x+y}$: Results screen}
\end{figure}

>>>>

\section{Features Extraction}
Our main goal is to develop a system that compliments the already available search capabilities in Invenio and in particular the CDS instance. Our system was developed on top of the  the Apache Solr/Lucene framework. CDS already uses Solr as a search engine layer, so developing $5e^{x+y}$(CDS) using this technology supposed a reasonable choice. One of the main steps while building an IR tool, is understanding the data and what kind of features can be extracted from it: If the system is based on text, then tasks such as stemming, handling of synonyms, misspelling and so on, should be taken into consideration; if the system being developed is based on images, then examples of features can be histograms of colors, geometrical features and so on. During the literature review, we discovered that most of the projects handled the notational aspects of the expressions, that is, the real naming of the terms in a given equation and each project developed its own set of heuristics to handle the structural component.

\subsection{Notational Features}
Notational features correspond to the elements that relate to the actual naming of the variables, constants and numeric quantities found in an equation. Here we extract tokens that represent single leaf nodes (E.g. <mi>, <mn>, <mo> tags), and simple constructs like subscript and superscript elements, fractions, roots.
We also extract bi-grams and tri-grams as an analogy to what is done for text indexing in current systems.

@@

\subsubsection{Unicode Normalizing} 
The first issue we discovered is that because of the wide range of different characters to represent the possible mathematical symbols, there is need for an automatic way to match different possible variations of a single character. An example of this situation is the specification of an natural number, sometimes denoted by the letter N. In this case, at least three different common representations were identified: $ N = 1 $ (Using character with unicode code point 0x004e), $\mathcal{N} = 1$ (0xd835 0xdc41 or in latex expressed as \lstinline|\mathcal{N}|) and $\mathbb{N} = 1$ (0x2115 or in latex expressed as \lstinline|\mathbb{N}|). \\
  To address this situation, we rely on the Unicode equivalence framework. In the unicode specification, sometimes the same character has two or more different code-points because of backwards compatibility with other encoding standards or because the same character has different essential meanings. Another group of equal characters with different code-points are pre-combined characters such as the Latin letter \~{N} which has Unicode point 0x00D1 and also is the sequence of code-points 0x004E (Latin letter N) and 0x0303 (the combining character tilde). Taking this sample into account, Unicode defines the character composition and decomposition operations as replacing the decomposed representation with the pre-composed one and vice-versa. Finally, Unicode defines two types of equivalences between characters: Canonical and compatible. Canonical equivalent characters are assumed to have the same appearance and meaning. Compatible ones are allowed to have slightly different graphical representations but the same meaning in some contexts. One example of compatible equivalent characters is the Roman number 12 {\unicodefontⅫ} with code-point 0x216b and the sequence of characters XII with codepoints 0x0058 0x0049 0x0049. Taking into account the composition or decomposition of characters and the canonical or compatible equivalences, Unicode specifies 4 different forms of a given character:
  \begin{itemize}
  \item NFD: Normalization Form Canonical Decomposition
  \item NFC: Normalization Form Canonical Composition
  \item NFKD: Normalization Form Compatibility Decomposition
  \item NFKC: Normalization Form Compatibility Composition
  \end{itemize}

@@

Since our goal is to match as many similar characters as possible, our system employs the NFKD representation of a character. For a given token, we test whether if it is already equal to its NFKD. If not, for all the characters that are not combining nor control characters, we add the token into the same position as the original one.
For example, given the token <mi>\r{A}</mi>, this will lead to the sequence <mi>\r{A}</mi><mi>A</mi> and the token <mi>{\unicodefontⅫ}</mi> will produce <mi>{\unicodefontⅫ}</mi><mi>X</mi><mi>I</mi><mi>I</mi>.

\subsubsection{Operator grouping}
Even though the Unicode normalization step works for a big part of the cases presented in table \ref{unicode_characters}, there are still some groups of characters which would not be matched (even partially), such as some types of integrals. To address these groups, a precomputed table was created for some general groups of operators which have some similar meaning. For each token that belongs to some of this predefined groups, a second token is created identifying the group.
For example the token <mo>{\unicodefont ∮}</mo> is expanded into <mo>{\unicodefont ∮}</mo><mo>INTEGRALS</mo>, similarly the token <mo>{\unicodefont ∫}</mo> is expanded into <mo>{\unicodefont ∫}</mo><mo>INTEGRALS</mo> and then both will have a partial match in the INTEGRALS token. 
The Unicode specification does not define a strict rule to group similar characters based on their semantics. For general operators, we used the grouping provided by Xah Lee in his website\cite{math_groups}.

@@

\subsubsection{Numeric Approximation} 
Some equations relate to specific numeric quantities. While most of the previously explorer systems handled <mn> tags by only representing them as a constant identifier, this approach is not suitable. A lot of articles in CDS relate to specific experiments under certain conditions, and require a high level of precision in the measurements. However, different experiments of the same phenomenon, can result in different measurements and we wanted to include this variability in our system. To handle this, when a token with a certain numeric quantity is discovered, we apply iterative rounding and index all the different approximations in the same position as the original element. For example the token <mn>2.0135</mn> will produce the following sequence of tokens: <mn>2.0135</mn><mn>2.014</mn><mn>2.01</mn><mn>2.0</mn>.

\subsection{Structural Features}
The structure of a mathematical expression compliments the information that can be obtained over generic substructures where the naming of the variables is irrelevant.

\subsubsection{Algebraic structures}
The first set of structures consists of purely algebraic relationships between the expressions.
Inspired by predefined integral tables for different kinds of expressions, we decided to extract, from a given equation, whether it has an occurrence of these kinds of patterns or not.
An example of this type of structures would be: $X\textunderscore*(X\textunderscore + A\textunderscore)$ where $X\textunderscore$ and $A\textunderscore$ can be any type of expression. Sample equations matching this pattern can be $C (C+1)$ or $\frac{(\sin{\theta}\cos{\theta})^2}{2a} (\frac{(\sin{\theta}\cos{\theta})^2}{2a} + \tan{\frac{\theta+\phi}{2}})$. This structural analysis of equations is more powerful than the unification of variables and constants proposed in previously discussed projects because these patterns, as shown in the second example, can match an arbitrary complex expression. The drawback of our solution is the need to manually select representative and useful patterns. This selection was done based on standard based tables of integrals similar to the ones that CASs employ to compute them.

\subsubsection{Subscript/Superscript variations}
This set of structures represent common operations that can appear in a given expression, but its identification is based on notation factors. A simple example of this can be the matrix decomposition $A = VDV^{-1}$ where the actual name of the variables $A$ and $V$ is not meaningful but the real importance is the fact that a given variable appears alongside its inverse in a matrix multiplication form. We would like to assign a partial match to any other kind of this decomposition say $B = MRM^{-1}$. These types of structures live in-between the notational and structural categories but since we identify them using Mathematica, we encompass them as structural features. For this type of structures, we identify simple operations between an element and some possible variations of the same element with some subscript/superscript.

@@

Additionally, there are common patterns to specific domains which can improve the quality of the retrieved results. Some specific patterns we identified include: $originalParticle \rightarrow subparticle1, subparticle2, subparticle3$ (for representing particle decaying processes) $S = \int^{t2}_{t1} \mathcal{L} dt$ (for an action in terms of its Lagrangian) and $SU(n)$ (for describing special unitary groups). The main drawback of this approach is the need to manually detect, implement and maintain long lists of patterns and for this reason we did not try to assemble an exhaustive list.

@@

The Lucene framework defines some key concepts which ultimately are mapped into concrete classes. A {\codefont Document} is the minimum retrieval unit. For $5e^{x+y}$ we could either use a complete article as a document or a single equation. We choose the latter because it makes it easier to focus on extracting features from a single equation. A document contains a set of {\codefont fields} which represent different types of information that it can store. Each field has different options such as whether it would be indexed and/or stored. A field can also indicate whether it will be a {\codefont Term Vector}. Our fields are as follow: 
\begin{itemize}
\item math\_notational\_field: This field stores the notational features as a Term Vector. 
\item math\_structural\_field: This field stores the structural features as a Term Vector. 
\item math\_normalized\_notational\_field: This field stores the notational features of the normalized expression as a Term Vector. 
\item filename: This is an auxiliary field which stores the CDS record / article name where the equation was found.
\item number\_occurrences: Another auxiliary field that is used to score final results.
\end{itemize}

Each field is associated to a particular {\codefont Analyzer}. An analyzer is a specific class that handles an element from a field and groups all the operations that are to be performed. For this purpose we implemented {\codefont SolrNotationalAnalyzer}, {\codefont SolrStructuralAnalyzer} and {\codefont SolrNormalizedNotationalAnalyzer}. 
The Solr prefix is added to clarify the fact that the class extends a {\codefont SolrAnalyzer} which extends a normal lucene {\codefont Analyzer}.

@@

\begin{itemize}
\item {\codefont MultiplePatternTokenizer}: This class is in charge of applying a set of regular expressions over the given expression in order to extract the single elements that were presented. Lucene incorporates the class {\codefont PatternTokenizer} but it only accepts one regular expression.
\item {\codefont UnicodeNormalizingFilter}: This class receives single tokens and verifies if the token is in Unicode NFKD. In negative case, it emits an additional token in the normalized form.
\item {\codefont SynonymExpandingFilter}: This class verifies if the received token belongs to a predefined category of elements and emits an additional token with it.
\item {\codefont NumericalRoundingFilter}: This class processes {\codefont <mn>} tokens. When a number contains a decimal part, it applies the standard rounding mode iteratively and for each one, it emits a token for each step. 
\end{itemize}

@@

\subsubsection{Error handling}
Sometimes, after certain time, the connection between {\codefont KernelLink} and {\codefont Mathematica} would be lost launching an exception and all following invocations would fail as well.
Also, re-instantiating a KernelLink would yield an Exception. For this scenario, killing the associated processes to Mathematica was necessary with {\codefont killall -s 9 Mathematica; killall -s 9 MathKernel
} or {\codefont taskkill /im Mathematica.exe; taskkill /im MathKernel.exe} depending on the OS the system is running on.

@@

Mathematica took around 10 minutes to interpret it.
In the ideal case we would be able to identify the malformed expressions and pass to Mathematica only well-formed equations to be processed but as this simple example shows, that is a very challenging task and may require more work.
As a remedial situation, we used a fixed timeout to stop the processing of such expressions.

@@

A demo interface over the {\codefont pu} branch was developed to showcase the system, however to have a complete integration with CDS, more work is required. The demo setup employs the same dataset as used for experimentation.

\begin{figure}
\includegraphics[height=8 cm]{figures/input_interface.png}
\label{input_int}
\caption{Web interface of $5e^{x+y}$: Query input screen}
\end{figure}

\begin{figure}
\includegraphics[height=8 cm]{figures/results_interface.png}
\label{results_int}
\caption{Web interface of $5e^{x+y}$: Results screen}
\end{figure}

### 2014-03-10 docs/thesis/chap-introduction.tex
wrote clusring results work on document
proofreading fixes

\subsection{CDS Invenio}
CDS\cite{CDS1} (CERN Document Server) Invenio is the integrated digital library system developed and used at CERN\cite{CDS2}. To the date, it contains more than one million records and more than 400000 full-text documents. It provides tools to manage the complete workflow of a document taking care of the submission, annotation, edition, storage, searching, retrieval, displaying among other phases. Figure \ref{cds_screenshot} presents the search results for a simple query in CDS.

@@

Invenio\cite{invenio} is the software running behind CDS and is an open source project developed in parallel to CDS at CERN. Besides CDS, Invenio also supports around thirty scientific institutions worldwide including Fermilab, SLAC National Laboratory, INSPIRE and EPFL. 
Invenio is composed of several modules where each one can be mapped to a specific step in the workflow of a record. Figure \ref{invenio_architecture} presents the global architecture of Invenio. A detailed explanation of each module can be accessed in\cite{invenio_modules}

@@

\section{Project overview}
\subsection{Motivation}
The initial statement for this project was more generic and was encompassed as New ways to programmatically, accurately and efficiently extract data and metadata from digital files. During a first exploration around this topic, we focused our attention to the mathematical content that is stored in these files. After more discussions we identified that the extraction, storage, indexing and finally searching for mathematical content would be an and very useful project for CDS and an interesting research direction. CDS groups records in different categories or Collections such as Published Articles, Preprint, Photos, Books, General Talks among others. Some of these collections contain scientific documents in all of the different research areas where CERN is involved, and therefore a big amount of mathematical expressions are contained there. Only the Preprints collection contains 698581 records to the date, which harvest documents from services like ArXiV\cite{arxiv} where most of the documents are in the areas of physics, mathematics and statistics which are very rich in mathematical content.
Currently Invenio provides different ways of searching for records by specifying metadata fields like author:"Albert Einstein" or by keywords. However, there is no way to search for a given mathematical expression. The current workarounds would be to try to search for the name of the given expression if it is common enough to have been named like \emph{Schr\"oodinger equation}. This approach most of the times is not enough since most of the equations are not named and even named ones can be rewritten in several ways and each one may have a different importance to the user. This combination of factors, motivates the development of a complete system that allows users to search for relevant documents, based on mathematical expressions.

@@

\begin{itemize}
\item Explore ways to automatically extract the mathematical content from a document collection
\item Investigate different approaches and formats to store mathematical expressions
\item Identify relevant features in a mathematical expression and how to use them to index equations
\item Using the created index, propose ways to efficiently retrieve relevant expressions and articles given an input query
\item Implement the previous steps into a complete system (A search engine) and integrate it into the Invenio software
\item Evaluate different approaches and the quality of the provided results
\item Identify deficiencies in our work and propose solutions and further direction on work
\end{itemize}

>>>>

\subsection{CDS Invenio}
CDS\cite{CDS1} (CERN Document Server) Invenio is the integrated digital library system developed and used at CERN\cite{CDS2}. To the date, it contains more than one million records and more than 400000 full-text documents. It provides tools to manage the complete workflow of a document taking care of the submission, annotation, editing, storage, searching, retrieval, displaying among other phases. Figure \ref{cds_screenshot} presents the search results for a simple query in CDS.

@@

Invenio\cite{invenio} is the software running behind CDS and is an open source project developed in parallel to CDS at CERN. Besides CDS, Invenio also supports around thirty scientific institutions worldwide including Fermilab, SLAC National Laboratory, INSPIRE and EPFL. 
Invenio is composed of several modules, each one of which can be mapped to a specific step in the workflow of a record. Figure \ref{invenio_architecture} presents the global architecture of Invenio. A detailed explanation of each module can be accessed in \cite{invenio_modules}

@@

\section{Project overview}
\subsection{Motivation}
The initial statement for this project was more generic and was encompassed as "New ways to programmatically, accurately and efficiently extract data and metadata from digital files". During a first exploration around this topic, we focused our attention to the mathematical content that is stored in these files. After more discussions we identified that the extraction, storage, indexing and finally searching for mathematical content would be a very useful project for CDS and an interesting research direction. CDS groups records in different categories or Collections such as Published Articles, Preprint, Photos, Books, General Talks among others. Some of these collections contain scientific documents in all of the different research areas where CERN is involved, and therefore a big amount of mathematical expressions are contained there. Only the Preprints collection contains 698581 records to date, harvesting documents from services like ArXiV\cite{arxiv} where most of the documents are in the areas of physics, mathematics and statistics which are very rich in mathematical content.
Currently Invenio provides different ways of searching for records by specifying metadata fields like author:"Albert Einstein" or by keywords. However, there is no way to search for a given mathematical expression. The current workarounds would be to try to search for the name of the given expression if it is common enough to have been named like \emph{Schr\"oodinger equation}. This approach most of the times is not enough since most of the equations are not named and even named ones can be rewritten in several ways and each one may have a different importance to the user. This combination of factors, motivates the development of a complete system that allows users to search for relevant documents, based on mathematical expressions.

@@

\begin{itemize}
\item Explore ways to automatically extract the mathematical content from a document collection
\item Investigate different approaches and formats to store mathematical expressions
\item Identify relevant features in a mathematical expression and how to use them to index equations
\item Using the created index, propose ways to efficiently retrieve relevant expressions and articles given an input query
\item Implement the previous steps into a complete system (a search engine) and integrate it into the Invenio software
\item Evaluate different approaches and the quality of the provided results
\item Identify deficiencies in our work and propose solutions and further directions
\end{itemize}

### 2014-03-10 docs/thesis/chap-theoretical_background.tex
cleaned useless docs
proofreading fixes

\chapter{Theoretical Background}
\label{chapter-theoretical_background}
Information Retrieval (IR) consists on the activity of identifying relevant documents from a collection based on some input parameters.
Different models have been developed through the evolution of the field. In the subsequent sections, the most relevant to this work are presented.

@@

>>>>

\chapter{Theoretical Background}
\label{chapter-theoretical_background}
Information Retrieval (IR) consists of the activity of identifying relevant documents from a collection based on some input parameters.
Different models have been developed through the evolution of the field. In the subsequent sections, the most relevant models to this work are presented.

@@

\section{Latent Dirichlet Allocation}
